---
title: "PV"
author: "Yves"
date: "26 December 2024"
output:   
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
---

```{r setup, include=FALSE}
  library(shinyWidgets)

  RV <- reactiveValues()

  source("C:\\Users\\ylace\\OneDrive\\Desktop\\STUFF\\PVGIS\\Model.R")
  defaults <- defaults()
  RV$input <- defaults
```

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=100}
-------------------------------------
### <b></b>

```{r, echo=FALSE}
  actionButton("reset1", "Map")
  HTML("<br><br>")
  actionButton("reset2", "Home")
```

```{r, echo=FALSE}
observeEvent(input$reset1,{
    lat <- RV$input$location['lat',1]
    lon <- RV$input$location['lon',1]
    print(paste("reset the map", lat, lon))
    
    proxy <- leafletProxy("EC")
    proxy %>% clearShapes() %>% 
     addProviderTiles("CartoDB.Positron") %>% 
     setView(lng=lon, lat=lat, zoom = 12) %>% 
     addCircles(lng=lon, lat=lat, radius=500, color='red')
})
```

```{r, echo=FALSE}
observeEvent(input$reset2,{
    lat <- defaults$location['lat', 1]
    lon <- defaults$location['lon', 1]
    print(paste("relocate the map to its original location", lat, lon))
    proxy <- leafletProxy("EC")
    proxy %>% clearShapes() %>% 
     addProviderTiles("CartoDB.Positron") %>% 
     setView(lng=lon, lat=lat, zoom = 12) %>% 
     addCircles(lng=lon, lat=lat, radius=500, color='red')
})
```

```{r, echo=FALSE}
  observeEvent(input$EC_click, {
    lat <- input$EC_click$lat
    lon <- input$EC_click$lng
    RV$input$location['lat',1] <- lat
    RV$input$location['lon',1] <- lon
    print(paste("update location after clicking on the map",
                RV$input$location['lat',1], RV$input$location['lon',1]))

    proxy <- leafletProxy("EC")
    proxy %>% clearShapes() %>% 
      addProviderTiles("CartoDB.Positron") %>% 
      setView(lng=lon, lat=lat, zoom = 12) %>% 
      addCircles(lng=lon, lat=lat, color='red',radius = 500) 
})
```

Column {}
-------------------------------------

### <b>Solar radiation and photovoltaic system performance for any location in the world, except the North and South Poles.</b>
```{r, echo=FALSE}
output$EC <- renderLeaflet({
    lat <- RV$input$location['lat',1]
    lon <- RV$input$location['lon',1]
    print(paste("set up the map", lon, lat))
   
    leaflet() %>% 
      addProviderTiles("CartoDB.Positron") %>% 
      setView(lng=lon, lat=lat, zoom = 12) %>% 
      addCircles(lng=lon, lat=lat, radius=500, color='red')
})
leafletOutput('EC')
```

Column {}
-------------------------------------
### <b>Location</b>
<b>Click on the map to change location</b>

```{r, echo=FALSE}
  renderUI({
  HTML(RV$input$location %>% kbl() %>% kable_styling())
})
```

###

Adjust input parameters on the next page. 

Once you are done, click this button to update the parameters.

```{r, echo=FALSE}
  actionButton("launch", "Update")
```

```{r, echo=FALSE}
observeEvent(input$launch,{

  RV$input$system["angle", 1] <- input$s1
  RV$input$system["aspect", 1] <- input$s2
  RV$input$system["pvtechchoice", 1] <- input$s3
  RV$input$system["mountingplace", 1] <- input$s4
  RV$input$system["loss", 1] <- input$s5
  RV$input$system["peakpower", 1] <- input$s6
  RV$input$system["trackingtype", 1] <- input$s7
  
  RV$input$off_grid["batterysize",1] <- input$o1
  RV$input$off_grid["cutoff",1] <- input$o2
  RV$input$off_grid["consumptionday",1] <- input$o3
  
  RV$input$calculations["optimalangles",1] <- as.numeric(input$c1)
  RV$input$calculations["pvcalculation",1] <- as.numeric(input$c2)
  RV$input$calculations["components",1] <- as.numeric(input$c3)
  
})
```

###

Photovoltaic Geographical Information System (PVGIS) 

  - online tool ... https://joint-research-centre.ec.europa.eu/photovoltaic-geographical-information-system-pvgis_en
  - API ... https://joint-research-centre.ec.europa.eu/photovoltaic-geographical-information-system-pvgis/getting-started-pvgis/api-non-interactive-service_en
  - entry point ... https://re.jrc.ec.europa.eu/api/v5_1/tool_name?param1=value1&param2=value2&...

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=300}
-------------------------------------
### <b>System</b>
```{r, echo=FALSE}
  renderUI({
    df <- data.frame(cbind(name=rownames(defaults$system), desc=defaults$system$desc))
    HTML(df %>% kbl() %>% kable_styling())
  })
```


Column {data-width=150}
-------------------------------------
### <b>System</b>
```{r, echo=FALSE}
  textInput("s1",rownames(defaults$system)[1], value=defaults$system[1,1])  
  textInput("s2",rownames(defaults$system)[2], value=defaults$system[2,1])  
  textInput("s3",rownames(defaults$system)[3], value=defaults$system[3,1])  
  textInput("s4",rownames(defaults$system)[4], value=defaults$system[4,1])  
  textInput("s5",rownames(defaults$system)[5], value=defaults$system[5,1])  
  textInput("s6",rownames(defaults$system)[6], value=defaults$system[6,1])  
  textInput("s7",rownames(defaults$system)[7], value=defaults$system[7,1])  

```

Column {data-width=300}
-------------------------------------
### <b>Off grid</b>
```{r, echo=FALSE}
  renderUI({
    df <- data.frame(cbind(name=rownames(defaults$off_grid), desc=defaults$off_grid$desc))
    HTML(df %>% kbl() %>% kable_styling())
  })
```

### <b>Calculations</b>
```{r, echo=FALSE}
  renderUI({
    df <- data.frame(cbind(name=rownames(defaults$calculations), desc=defaults$calculations$desc))
    HTML(df %>% kbl() %>% kable_styling())
  })
```

Column {data-width=150}
-------------------------------------
### <b>Off grid</b>
```{r, echo=FALSE}
  textInput("o1",rownames(defaults$off_grid)[1], value=defaults$off_grid[1,1])  
  textInput("o2",rownames(defaults$off_grid)[2], value=defaults$off_grid[2,1])  
  textInput("o3",rownames(defaults$off_grid)[3], value=defaults$off_grid[3,1])  
```

### <b>Calculations</b>
```{r, echo=FALSE}
  switchInput(inputId = "c1", value = TRUE,label=rownames(defaults$calculations)[1])  
  switchInput(inputId = "c2", value = TRUE,label=rownames(defaults$calculations)[2])  
  switchInput(inputId = "c3", value = TRUE,label=rownames(defaults$calculations)[3])  
```


Column {data-width=200}
-------------------------------------
### <b>Inputs</b>
```{r, echo=FALSE}
  renderUI({
    df <- data.frame(rbind(
      c(value=input$s1, label=rownames(defaults$system)[1]),  
      c(value=input$s2, label=rownames(defaults$system)[2]),  
      c(value=input$s3, label=rownames(defaults$system)[3]),  
      c(value=input$s4, label=rownames(defaults$system)[4]),  
      c(value=input$s5, label=rownames(defaults$system)[5]),  
      c(value=input$s6, label=rownames(defaults$system)[6]),  
      c(value=input$s7, label=rownames(defaults$system)[7]),  
  
      c(value=input$o1, label=rownames(defaults$off_grid)[1]),  
      c(value=input$o2, label=rownames(defaults$off_grid)[2]), 
      c(value=input$o3, label=rownames(defaults$off_grid)[3]),
    
      c(value=as.numeric(input$c1), label=rownames(defaults$calculations)[1]),  
      c(value=as.numeric(input$c2), label=rownames(defaults$calculations)[2]), 
      c(value=as.numeric(input$c3), label=rownames(defaults$calculations)[3])
    ))
    HTML(df %>% kbl() %>% kable_styling())
  })
```


<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=200}
-------------------------------------
### <b>TMY</b>

Typical meteorological year (TMY) is a set of meteorological data 
for every hour in a year for a given geographical location.

```{r, echo=FALSE}
actionButton("TMY", "")
```

Column {}
-------------------------------------
### 

```{r, echo=FALSE}
observeEvent(input$TMY, {
  print("processing TMY")
  source("C://Users//ylace//OneDrive//Desktop//STUFF//PVGIS//tmy.R")
  q <- paste0(
    url=url,
    "tmy?",
    "lat=", RV$input$location['lat',1], 
    "&lon=", RV$input$location['lon',1],
    "&outputformat=", outputformat
  )
  RV$TMY <- TMY(q)
})
```

```{r, echo=FALSE}
renderUI({
  HTML(RV$TMY$meta %>% kbl() %>% kable_styling())
})
```

###
```{r, echo=FALSE}
renderPlotly({

  if ( ! is.null(RV$TMY) ) {
    plot_ly(type='scatter', mode='markers', showlegend=TRUE) %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~T2m, name="T2m") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~RH, name="RH") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="G.h.") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="Gb.n.") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="Gd.h.") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="IR.h.") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="WS10m") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="WD10m") %>%
      add_trace(data=RV$TMY$tmy, x=~time.UTC., y=~G.h., name="SP") %>%
      layout(yaxis=list(title="Value"))
  }
})
```

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=300}
-------------------------------------
### <b>Hourly Radiation</b>

Solar irradiance and PV output data for every hour in a multi-year period.

```{r, echo=FALSE}
actionButton("HR", "")
```

### <b>Optimal panel geometry</b>
```{r, echo=FALSE}
renderUI({
  if ( ! is.null(RV$Hourly_radiation) ) {
    if ( RV$Hourly_radiation$input$value[12] == "TRUE" | 
       RV$Hourly_radiation$input$value[14] == "TRUE") {
        HTML(RV$Hourly_radiation$input[c(11,13),] %>% kbl() %>% kable_styling()) 
    }
  } 
})
```

Column {}
-------------------------------------
### 

```{r, echo=FALSE}
observeEvent(input$HR, {
  print("processing Hourly Radiation")
  source("C://Users//ylace//OneDrive//Desktop//STUFF//PVGIS//hourly_radiation.R")
  q <- paste0(
    url=url,
    "seriescalc?",
    "lat=", RV$input$location['lat',1], 
    "&lon=", RV$input$location['lon',1],
    "&angle=", RV$input$system['angle',1],
    "&aspect=", RV$input$system['aspect',1],
    "&loss=", RV$input$system['loss', 1],
    "&peakpower=", RV$input$system['peakpower', 1],
    "&optimalangles=", RV$input$calculations['optimalangles',1],
    "&pvcalculation=", RV$input$calculations['pvcalculation',1],
    "&components=", RV$input$calculations['components',1],
    "&startyear=", 2023,
    "&outputformat=", outputformat
  )
  RV$Hourly_radiation <- Hourly_radiation(q)
})
```

```{r, echo=FALSE}
renderUI({
  HTML(RV$Hourly_radiation$meta %>% kbl() %>% kable_styling())
})
```

###
```{r, echo=FALSE}
renderPlotly({

  if ( ! is.null(RV$Hourly_radiation) ) {
plot_ly(type='scatter', mode='lines', showlegend=TRUE) %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~Gb.i., name="Gb.i.") %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~Gd.i., name="Gd.i.") %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~Gr.i., name="Gr.i.") %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~H_sun, name="H_sun") %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~T2m, name="T2m") %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~WS10m, name="WS10m") %>%
  add_trace(data=RV$Hourly_radiation$hourly, x=~time, y=~P, name="P")
  }
})
```

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=300}
-------------------------------------
### <b>Grid Connected</b>

Monthly average values of PV system energy output.

In-plane irradiation per month and for the full year.

```{r, echo=FALSE}
actionButton("GC", "")
```

### <b>Optimal panel geometry</b>
```{r, echo=FALSE}
renderUI({
  if ( ! is.null(RV$Grid_connected) ) {
    if ( RV$Grid_connected$input$value[11] == "TRUE" | 
       RV$Grid_connected$input$value[13] == "TRUE") {
        HTML(RV$Grid_connected$input[c(10,12),] %>% kbl() %>% kable_styling()) 
    }
  } 
})
```

Column {}
-------------------------------------
### 

```{r, echo=FALSE}
observeEvent(input$GC, {
  print("processing Grid Connected")
  source("C://Users//ylace//OneDrive//Desktop//STUFF//PVGIS//grid_connected.R")
  q <- paste0(
    url=url,
    "PVcalc?",
    "lat=", RV$input$location['lat',1], 
    "&lon=", RV$input$location['lon',1],
    "&pvtechchoice=", RV$input$system['pvtechchoice',1],
    "&mountingplace=", RV$input$system['mountingplace',1],
    "&angle=", RV$input$system['angle',1],
    "&aspect=", RV$input$system['aspect',1],
    "&peakpower=", RV$input$system['peakpower',1],
    "&loss=", RV$input$system['loss',1],
    "&optimalangles=", RV$input$calculations['optimalangles',1],
    "&outputformat=", outputformat
  )
  RV$Grid_connected <- Grid_connected(q)
})
```

```{r, echo=FALSE}
renderUI({
  HTML(RV$Grid_connected$meta %>% kbl() %>% kable_styling())
})
```

###
```{r, echo=FALSE}
renderPlotly({
  if ( ! is.null(RV$Grid_connected) ) {
    plot_ly(type='scatter', mode='lines', showlegend=TRUE) %>%
      add_trace(data=RV$Grid_connected$out, x=~month, y=~E_d, name="E_d") %>%
      add_trace(data=RV$Grid_connected$out, x=~month, y=~E_m, name="E_m") %>%
      add_trace(data=RV$Grid_connected$out, x=~month, y=~H.i._d, name="H.i._d") %>%
      add_trace(data=RV$Grid_connected$out, x=~month, y=~H.i._m, name="H.i._m") %>%
      add_trace(data=RV$Grid_connected$out, x=~month, y=~SD_m, name="SD_m")
  }
})
```

###
```{r, echo=FALSE}
renderUI({
  HTML(RV$Grid_connected$totals %>% kbl() %>% kable_styling())
})
```

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=200}
-------------------------------------
### <b>Horizon Profile</b>

List of horizon height values for the given location. 

List of sun elevation angles at the June and December solstice. 

The horizon height is given at 48 equidistant directions starting 
from north and moving clockwise towards east, south and then west. 

The sun elevation angles are given at half-hour intervals starting at 00:00UTC.


```{r, echo=FALSE}
actionButton("HP", "")
```

Column {data-width=300}
-------------------------------------
### 

```{r, echo=FALSE}
observeEvent(input$HP, {
  print("processing Horizon_Profile")
  source("C://Users//ylace//OneDrive//Desktop//STUFF//PVGIS//horizon_profile.R")
  q <- paste0(
    url=url,
    "printhorizon?",
    "lat=", RV$input$location['lat',1], 
    "&lon=", RV$input$location['lon',1],
    "&outputformat=", outputformat
  )
  RV$Horizon_profile <- Horizon_Profile(q)
})
```

```{r, echo=FALSE}
renderUI({
  HTML(RV$Horizon_profile$meta %>% kbl() %>% kable_styling())
})
```

Column {}
-------------------------------------
### 
```{r, echo=FALSE}
renderPlotly({
  if ( ! is.null(RV$Horizon_profile) ) {
    plot_ly(type='scatter', mode='lines', showlegend=TRUE) %>%
      add_trace(data=RV$Horizon_profile$horizon_profile, x=~A, y=~H_hor, name="horizon_profile") %>%
      add_trace(data=RV$Horizon_profile$winter_solstice, x=~A_sun, y=~H_sun, name="winter_solstice") %>%
      add_trace(data=RV$Horizon_profile$summer_solstice, x=~A_sun, y=~H_sun, name="summer_solstice")
  }
})
```

### <b>Horizon profile</b>
```{r, echo=FALSE}
renderLeaflet({
  if ( ! is.null(RV$Horizon_profile) ) {
    
    point <- as.numeric(c(RV$Horizon_profile$input$Value[2], RV$Horizon_profile$input$Value[1]))

    Path.s <- data.frame(
      name="summer_solstice", RV$Horizon_profile$summer_solstice, 
      destPoint(p=point, b=RV$Horizon_profile$summer_solstice$A_sun, d=100))
      
    
    colnames(Path.s) <- c("name", "A", "H", "lon", "lat")
    Path.w <- data.frame(
      name="winter_solstice", RV$Horizon_profile$winter_solstice, 
      destPoint(p=point, b=RV$Horizon_profile$winter_solstice$A_sun, d=200))
    colnames(Path.w) <- c("name", "A", "H", "lon", "lat")
    Path.p <- data.frame(
      name="horizon_profile", RV$Horizon_profile$horizon_profile, 
      destPoint(p=point, b=RV$Horizon_profile$horizon_profile$A, d=300))
    colnames(Path.p) <- c("name", "A", "H", "lon", "lat")

    Path <- rbind(Path.s, Path.w)
    pal   <- colorNumeric(palette = "Spectral", domain = Path$H)
    pal.p <- colorNumeric(palette = "Spectral", domain = Path.p$H)

    leaflet() %>% addTiles() %>% 
      setView(lng=point[1], lat=point[2], zoom = 15) %>%
      addCircles(data=Path.p, lng=~lon, lat=~lat, 
                 radius=5, color=~pal.p(H),
                 label=~paste(name, A, H))  %>% 
      addLegend(position = "topright", pal=pal.p, values=Path.p$H)
  }
})
```

<b></b> {data-icon="ion-stats-bars"}
=====================================  

Column {data-width=200}
-------------------------------------
### <b>Off grid</b>

Monthly average values of PV system energy output.

Probability of battery charge reaching full or empty state. 

List of 10 values giving a histogram of battery charge state. 


```{r, echo=FALSE}
actionButton("OG", "")
```

Column {data-width=300}
-------------------------------------
### 

```{r, echo=FALSE}
observeEvent(input$OG, {
  print("processing Off_Grid")
  source("C://Users//ylace//OneDrive//Desktop//STUFF//PVGIS//off_grid.R")
  q <- paste0(
    url=url,
    "SHScalc?",
    "lat=", RV$input$location['lat',1], 
    "&lon=", RV$input$location['lon',1],
    "&peakpower=", RV$input$system['peakpower',1],
    "&angle=", RV$input$system['angle',1],
    "&aspect=", RV$input$system['aspect',1],
    "&batterysize=", RV$input$off_grid['batterysize',1],
    "&consumptionday=", RV$input$off_grid['consumptionday',1],
    "&cutoff=", RV$input$off_grid['cutoff',1],
    "&outputformat=", outputformat
  )
  RV$Off_grid <- Off_grid(q)
})
```

```{r, echo=FALSE}
renderUI({
  HTML(RV$Off_grid$meta %>% kbl() %>% kable_styling())
})
```

Column {}
-------------------------------------
### 
```{r, echo=FALSE}
renderPlotly({
  if ( ! is.null(RV$Off_grid) ) {
    plot_ly(type='scatter', mode='lines', showlegend=TRUE) %>%
      add_trace(data=RV$Off_grid$monthly, x=~month, y=~E_d, name="E_d") %>%
      add_trace(data=RV$Off_grid$monthly, x=~month, y=~E_lost_d, name="E_lost_d") %>%
      add_trace(data=RV$Off_grid$monthly, x=~month, y=~f_f, name="f_f") %>%
      add_trace(data=RV$Off_grid$monthly, x=~month, y=~f_e, name="f_e")
  }
})
```

### <b>Histogram</b>
```{r, echo=FALSE}
renderPlotly({
  if ( ! is.null(RV$Off_grid) ) {
    plot_ly(type='bar') %>%
      add_trace(data=RV$Off_grid$histogram, x=~f_CS, y=~CS_min, name="CS_min") %>%
      add_trace(data=RV$Off_grid$histogram, x=~f_CS, y=~CS_max, name="CS_max") %>%
      layout(yaxis=list(title="value"))
  }
})
```
